---
title: "Activity Monitoring"
author: "Marlon Gaviria Pérez"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    code_folding: show
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## <font color = 'blue'>*Read the Data*</font>

Please unfold the *code* bottom to see the respective lines of code

```{r read data,comment='',message=FALSE, warning=FALSE}
library(dplyr)

act <- read.csv('activity.csv')

total_steps <- na.omit(act %>% group_by(date) %>% summarise(total_steps=sum(steps)))

```

## <font color = 'blue'>*Histogram of the total number of steps taken each day*</font>

```{r histogram of total steps, warning=FALSE, message=FALSE}
total_steps <- na.omit(act %>% group_by(as.Date(date)) %>% summarise(total_steps=sum(steps)))
hist(total_steps$total_steps, main = 'Total steps each day',
     xlab = 'Steps',col = rainbow(n = 3,v = 0.5,alpha = 0.5))
grid()

```

## <font color = 'blue'>*Mean and median number of steps taken each day*</font>

```{r mean and meadian of steps, message=FALSE, warning=FALSE}
library(kableExtra)
med <- mean(total_steps$total_steps)
mad <- median(total_steps$total_steps)
mms <- data.frame(Values=c(med,mad), row.names=c('Mean','Median'))
kable(mms) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## <font color = 'blue'>*Time series plot of the average number of steps taken*</font>

```{r plot of the average of steps, message=FALSE, warning=FALSE}
avrg_steps <- na.omit(act %>% group_by(date=as.Date(date)) %>% summarise(avrg_steps=mean(steps)))
plot(avrg_steps$date,avrg_steps$avrg_steps, type = 'l',lwd = 5, xlab = 'Day', ylab = 'Steps', 
     col = 'cyan',)
points(avrg_steps$date,avrg_steps$avrg_steps, type = 'l', col = 'darkblue')
points(as.Date("2012-10-20"),40, cex=120,col = rgb(51/255,212/255,255/255,0.1),pch = 16)
grid()

```

## <font color = 'blue'>*The 5-minute interval that, on average, contains the maximum number of steps*</font>

```{r, max average interval, warning=FALSE,message=FALSE, comment=''}
act <- na.omit(transform(act,interval=as.character(interval)))
avgstps <- act %>% group_by(interval=as.character(interval)) %>% summarise(avrg_steps=mean(steps))
avgstps[which.max(avgstps$avrg_steps),1]
```

the interval with the maximun average of steps is (835-840) with 206 steps by average.

## <font color='blue'>*Code to describe and show a strategy for imputing missing data*</font>

```{r}
act_miss <- read.csv('activity.csv', sep = ',')

act_miss$steps[is.na(act_miss$steps)] <- mean(act_miss$steps,na.rm = T)
```

## <font color='blue'>*Histogram of the total number of steps taken each day after missing values are imputed*</font>

```{r host of total steps without NA, message=FALSE,warning=FALSE}
total_steps <- na.omit(act_miss %>% group_by(as.Date(date)) %>% summarise(total_steps=sum(steps)))
hist(total_steps$total_steps, main = 'Total steps each day',
     xlab = 'Steps',col = rainbow(n = 3,v = 0.5,alpha = 0.5))
grid()

```

## <font color='blue'>*Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends*</font>

```{r week and weekend avrg steps, warning=FALSE,message=FALSE}

library(lubridate)
library(dplyr)
weekend <- which( wday(as.Date(act_miss$date))==1 | wday(as.Date(act_miss$date))==7)

act_miss_weekend <- act_miss[weekend,]
act_miss_week <- act_miss[-weekend,]

act_week <- act_miss_week %>% group_by(interval = as.character(interval)) %>% summarise(avr=mean(steps))
act_weekend <- act_miss_weekend %>% group_by(interval = as.character(interval)) %>% summarise(avr=mean(steps))

act_week <- act_week[order(as.numeric(act_week$interval)),]
act_weekend <- act_weekend[order(as.numeric(act_weekend$interval)),]

par(mfrow = c(2,1))
# plot1
plot(act_week$interval,act_week$avr, type = 'l', xlab = 'Intervals', ylab = 'Average of steps',
     main='Week', col = 'cyan',lwd=4,fg='blue')
points(act_week$interval,act_week$avr, type = 'l', col = 'blue')
points(1250,75,pch=16,cex=100, col = rgb(51/255,212/255,255/255,0.1))
grid()
# plot2
plot(act_weekend$interval,act_weekend$avr, type = 'l', xlab = 'Intervals', ylab = 'Average of steps',
     main='Weekend', col = 'cyan',lwd=4,fg='blue')
points(act_weekend$interval,act_weekend$avr, type = 'l', col = 'blue')
points(1250,75,pch=16,cex=100, col = rgb(51/255,212/255,255/255,0.1))
grid()

```

